---
title: "Lake Acidity Modeling"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Modeling Acidity (pH levels) of bodies of water 
```{r}
library(faraway)
library(lmtest)
library(leaps)
acid_data = read.csv("lake_data_cleaned.csv")
```
```{r}
acid_data = read.csv("lake_data_cleaned.csv")
data.frame(acid_data)
```
##Multiple linear regression (collinearity, VIF,)
test out various hypotheses on single variable regressiioon, and multiple linear reg

##violations of assumptions (linearity, constant variation, normality of errors, interaction)

```{r}

acid_no3 = lm(PH_LAB..PH.Level. ~ NO3_UEQ_L..Nitrate., data=acid_data)
plot(PH_LAB..PH.Level. ~ NO3_UEQ_L..Nitrate., col="red", data=acid_data)
abline(acid_no3)

```
The above plot shows a linear model with pH level as the response variable and NO3 as the only predictor variable. It does not look like a good fit. We can explore the following violations:

linearity and constant variance
```{r}
plot(fitted(acid_no3), resid(acid_no3), col = "grey", pch = 20,
     xlab = "Fitted", ylab = "Residuals", main = "Fitted versus Residuals")
abline(h = 0, col = "darkorange", lwd = 2)
```
Clearly, we observe from the linear regression plot and Residuals vs Fitted plot that we violate linearity and constant variance.

```{r}
library(lmtest)
bptest(acid_no3)
```
Using the Breusch-Pagan test, we confirm that constant variance is violated because our p-value is very small. Therefore, we reject our null hypothesis, meaning the error does not have constant variance.

```{r}
par(mfrow=c(1,2))
hist(resid(acid_no3), xlab= "Residuals", main= "Histogram of Residuals", col= "blue")
qqnorm(resid(acid_no3), main = "Normal Q-Q Plot", col = "red")
qqline(resid(acid_no3), col = "green", lwd = 2)
```

##linear regression comparison (ANOVA) ISA

##polynomial regression ETHAN
test out various hypotheses on single variable regresion, and multiple linear reg

##violations of assumptions (linearity, constant variation, normality of errors, interaction, outliers (cooks)) ISA


###Variable Selection
     
#Using Backward Search
```{r}
acid_model = lm(PH_LAB..PH.Level. ~ . , data= acid_data)
backward_aic = step(acid_model, direction="backward", trace = 0)
backward_aic
```
```{r}
backward_bic = step(acid_model, direction="backward", k = log(length(resid(acid_model))), trace = 0)
backward_bic
```
     
Comparing the AIC and BIC Model with a level of $\alpha =  0.10$.
```{r}
anova(backward_bic, backward_aic)
```
Test Statistic: F = 4.139      
Distribution: On 2 P-Value = 0.01595.     
Decision: pvalue < alpha, we reject the null hypothesis and we choose the backward_aic model.
         
We can also compare the values of adjusted $R^2$ to figure out which is the "best" model. 
```{r}
summary(backward_aic)$adj.r.squared # 0.8030164
summary(backward_bic)$adj.r.squared # 0.802968
```
Comparing these $R^2$ model, we that the backward_aic model has a slightly higher adjusted $R^2$ value, so it will be the better model in this case. This is consistent with our anova test as well.     
     
#Using Forward Search: 
```{r}
acid_model_start = lm(PH_LAB..PH.Level. ~ 1 , data= acid_data)
forward_aic = step(acid_model_start, scope = PH_LAB..PH.Level. ~ ANC_UEQ_L..Acid.Neutralizing.Capacity. + CA_UEQ_L..Calcium. + CL_UEQ_L..Chloride. + COND_UM_CM..Conductivity.+K_UEQ_L..potassium.+ MG_UEQ_L..magnesium.+NA_UEQ_L..sodium.+ NH4_UEQ_L..Amonium.+NO3_UEQ_L..Nitrate.+SIO2_MG_L..silicon.dioxicde.+SO4_UEQ_L..sulfate.+WATERBODY_TYPE, direction="forward", trace = 0)
forward_aic
```
```{r}
forward_bic = step(acid_model_start, scope = PH_LAB..PH.Level. ~ ANC_UEQ_L..Acid.Neutralizing.Capacity. + CA_UEQ_L..Calcium. + CL_UEQ_L..Chloride. + COND_UM_CM..Conductivity.+K_UEQ_L..potassium.+ MG_UEQ_L..magnesium.+NA_UEQ_L..sodium.+ NH4_UEQ_L..Amonium.+NO3_UEQ_L..Nitrate.+SIO2_MG_L..silicon.dioxicde.+SO4_UEQ_L..sulfate.+WATERBODY_TYPE, direction="forward", k = log(length(resid(acid_model))), trace = 0)
forward_bic
```
        
Comparing the AIC and BIC Model with a level of $\alpha =  0.10$.
```{r}
anova(forward_bic, forward_aic)
```
Test Statistic: F = 4.139    
Distribution: On 2 P-Value = 0.01595   
Decision: pvalue < alpha, we reject the null hypothesis and we choose the forward_aic model.
        
We can also compare the values of adjusted $R^2$ to figure out which is the "best" model. 
```{r}
summary(forward_aic)$adj.r.squared # 0.8030164
summary(forward_bic)$adj.r.squared # 0.802968
```
Comparing these $R^2$ model, we that the forward_aic model has a slightly higher adjusted $R^2$ value, so it will be the better model in this case. This is consistent with the anova test. 

#Stepwise Search
```{r}
acid_model_start = lm(PH_LAB..PH.Level. ~ 1 , data= acid_data)
both_aic = step(acid_model_start, scope = PH_LAB..PH.Level. ~ ANC_UEQ_L..Acid.Neutralizing.Capacity. + CA_UEQ_L..Calcium. + CL_UEQ_L..Chloride. + COND_UM_CM..Conductivity.+K_UEQ_L..potassium.+ MG_UEQ_L..magnesium.+NA_UEQ_L..sodium.+ NH4_UEQ_L..Amonium.+NO3_UEQ_L..Nitrate.+SIO2_MG_L..silicon.dioxicde.+SO4_UEQ_L..sulfate.+WATERBODY_TYPE, direction="both", trace = 0)
both_aic
```
```{r}
both_bic = step(acid_model_start, scope = PH_LAB..PH.Level. ~ ANC_UEQ_L..Acid.Neutralizing.Capacity. + CA_UEQ_L..Calcium. + CL_UEQ_L..Chloride. + COND_UM_CM..Conductivity.+K_UEQ_L..potassium.+ MG_UEQ_L..magnesium.+NA_UEQ_L..sodium.+ NH4_UEQ_L..Amonium.+NO3_UEQ_L..Nitrate.+SIO2_MG_L..silicon.dioxicde.+SO4_UEQ_L..sulfate.+WATERBODY_TYPE, direction="both", k = log(length(resid(acid_model))), trace = 0)
both_bic
```
    
Using anova to compare these models with $\alpha= 0.10$
```{r}
anova(both_bic, both_aic)
```
Test Statistic: F = 4.139    
Distribution: On 2 P-Value = 0.01595   
Decision: pvalue < alpha, we reject the null hypothesis and we choose the both_aic model.
      
We can also compare $R^2$ values: 
```{r}
summary(both_aic)$adj.r.squared # 0.8030164
summary(both_bic)$adj.r.squared # 0.802968
```
The $R^2$ values for both_aic is slightly higher, thus a better model than the both_bic model. We choose the both_aic model. 
       
As we can see with both Backward, forward and stepwise search we ultimately find that the AIC model is better than the BIC model. Let's look at the violation of assumptions for both the AIC and BIC model using the results from the backward search (though regardless of the method we did, all three search methods ultimately had the same $R^2$ values and coefficients values.)
     
Let's find the model with the **best** AIC: 
```{r}
acid_model = lm(PH_LAB..PH.Level. ~ . , data= acid_data)
all_acid_model = summary(regsubsets(PH_LAB..PH.Level. ~ ., data = acid_data))
n = length(resid(acid_model))
p = length(coef(acid_model))
acid_mod_aic = n * log(all_acid_model$rss / n) + 2 * (2:p)
best_aic_ind = which.min(acid_mod_aic)
all_acid_model$which[best_aic_ind,]
```
From this, we can see that Acid Neutralizing Capacity, Calcium, Conductivity, Potassium, Magnesium, Sodium, Amonium, and Sulfate are the best predictors to use. 
         
```{r}
best_aic_model = lm(PH_LAB..PH.Level. ~ ANC_UEQ_L..Acid.Neutralizing.Capacity. + CA_UEQ_L..Calcium.+COND_UM_CM..Conductivity. + K_UEQ_L..potassium. + MG_UEQ_L..magnesium.+   NA_UEQ_L..sodium. + NH4_UEQ_L..Amonium. +    SO4_UEQ_L..sulfate. , data = acid_data)
summary(best_aic_model)$adj.r.squared # 0.8026063
```

```{r}
calc_loocv_rmse = function(model) {
  sqrt(mean((resid(model) / (1 - hatvalues(model))) ^ 2))
}

calc_loocv_rmse(best_aic_model) #0.3359839
calc_loocv_rmse(backward_aic) #0.3358015
```
We would want to use the model with the lowest RMSE, which in this case will be the backward_aic model. 



##polynomial regression comparison (ANOVA) ETHAN

##Transformations ch 14 ETHAN



