---
title: "Lake Acidity Modeling"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Modeling Acidity (pH levels) of bodies of water 


```{r}
acid_data = read.csv("lake_data_cleaned.csv")
data.frame(acid_data)
```
##Multiple linear regression
test out various hypotheses on single variable regressiioon, and multiple linear reg

##violations of assumptions (linearity, constant variation, normality of errors, interaction)

```{r}

acid_no3 = lm(PH_LAB..PH.Level. ~ NO3_UEQ_L..Nitrate., data=acid_data)
plot(PH_LAB..PH.Level. ~ NO3_UEQ_L..Nitrate., col="red", data=acid_data)
abline(acid_no3)

```
The above plot shows a linear model with pH level as the response variable and NO3 as the only predictor variable. It does not look like a good fit. We can explore the following violations:

linearity and constant variance
```{r}
plot(fitted(acid_no3), resid(acid_no3), col = "grey", pch = 20,
     xlab = "Fitted", ylab = "Residuals", main = "Fitted versus Residuals")
abline(h = 0, col = "darkorange", lwd = 2)
```
Clearly, we observe from the linear regression plot and Residuals vs Fitted plot that we violate linearity and constant variance.

```{r}
library(lmtest)
bptest(acid_no3)
```
Using the Breusch-Pagan test, we confirm that constant variance is violated because our p-value is very small. Therefore, we reject our null hypothesis, meaning the error does not have constant variance.

```{r}
par(mfrow=c(1,2))
hist(resid(acid_no3), xlab= "Residuals", main= "Histogram of Residuals", col= "blue")
qqnorm(resid(acid_no3), main = "Normal Q-Q Plot", col = "red")
qqline(resid(acid_no3), col = "green", lwd = 2)
```

##linear regression comparison (ANOVA) ISA

##polynomial regression ETHAN
test out various hypotheses on single variable regresion, and multiple linear reg

##violations of assumptions (linearity, constant variation, normality of errors, interaction, outliers (cooks)) ISA
##choose variables using aic, bic, adjusted r^2 CAITLYN

##polynomial regression comparison (ANOVA) ETHAN

##Transformations ch 14 ETHAN



